---
title: "Statistical Concepts"
---

```{r}
#| include: false
library(dplyr)
```

* For each variable, compute the following values. You can use the built-in functions or use the mathematical formulas.
    + expected value
    + variance
    + standard deviation

```{r}
random_vars <- readRDS("../data/random_vars.rds")
random_vars %>% summarise(expectedAge = mean(age), varianceAge = var(age), stdAge = sd(age))
```

```{r}
random_vars %>% summarise(expectedIncome = mean(income), varianceIncome = var(income), stdIncome = sd(income))
```

* Explain, if it makes sense to compare the standard deviations.

No!, because standard deviation only represents how far the samples are from the sample-mean. Whereas, covariance is the measure of linear dependency, but it can not be compared easily. Correlation can have value form -1 to 1 and can be used instead to interpret the relationship between two quantities.

* Then, examine the relationship between both variables and compute:
    + covariance
    + correlation

```{r}
random_vars %>% summarise(covariance = cov(age,income), correlation = cor(age,income))
```

* What measure is easier to interpret? Please discuss your interpretation.

As stated above, covariance is not easy interpret, therefore, correlation is easier to interpret.

* Compute the conditional expected value for:
    + E[income | age <= 18]
    + E[income | age $\epsilon$ [18,65)]
    + E[income | age >= 65]
    
```{r}
# E[income | age <= 18]
mean(random_vars %>% filter(age<=18) %>% select(income) %>% pull(income))
```

```{r}
# E[income | age ϵ [18,65)]
mean(random_vars %>% filter(age>=18 & age<65) %>% select(income) %>% pull(income))
```

```{r}
# E[income | age >= 65]
mean(random_vars %>% filter(age>=65) %>% select(income) %>% pull(income))
```